---
title: "Assignment 6: ODE & Sobol"
description: |
  This assignment analyzes model sensitivity with Sobol indicies. 
author: "Caroline Shepherd & Trevor Maggart"
output:
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
editor_options: 
  chunk_output_type: console
---

This code uses the model below to visualize forest growth and analyze how different parameters impact forest growth.

The model takes the following equations and parameters:

- dC/dt = r ∗ C for forests where C is below a threshold canopy closure
- dC/dt = g ∗ (1 − C/K) for forests where carbon is at or above the threshold canopy closure
- K is carrying capacity in units of carbon

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align='center')

library(tidyverse)
library(here)
library(sensitivity)
library(deSolve)
library(kableExtra)
```

```{r, forest growth function}
forest_growth = function(time, C, params) {
 
  if (C < params$canopy_closure) {
    dC = params$r * C
  } else {
    dC =  params$g * (1-(C/params$k))
  }
             
  return(list(dC)) 
}
```


```{r, initiate parameters & run ODE}
# start with an initial forest size (C) of 10
C_initial = 10

# set our parameter values
k = 250
r = 0.01
g = 2
canopy_closure = 50

# combine parameters into a dataframe using `cbind`
params <- cbind.data.frame(k, r, g, canopy_closure)

# create a sequence of years from 1-300
sim_length = seq(from = 1, to = 300)

# use our ODE solver from `deSolve` and save the result as a dataframe
result <- ode(y = C_initial, times = sim_length, func = forest_growth, parms = params)

# add column names
colnames(result) <- c("time", "C")

# save as dataframe
result <- as.data.frame(result)
```

```{r, plot the ODE results}
ggplot(result, aes(x = time, y = C)) + 
  theme_minimal() +
  geom_point(size = 0.4, color = "darkolivegreen") +
  labs(x = "Years",
       y = "Forest Size (kg C)") +
  scale_x_continuous(limits = c(0, 300), breaks = seq(0, 300, by=50), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 200), expand = c(0,0)) +
  theme(
    axis.title.x = element_text(size = 12, family = "Avenir", face = 'bold', vjust = -0.5),
    axis.title.y = element_text(size = 12, family = "Avenir", face = 'bold', vjust = 2)
  )
```
```{r}
# create a standard deviation variable of 10% to be used with all parameters
sd = 0.1

# create a sample size of 1000
n_sample = 1000

# create two sample sets of parameter values
## sample set 1
r_sobol = rnorm(mean = r, sd = r*sd, n = n_sample)
g_sobol = rnorm(mean = g, sd = g*sd, n = n_sample)
k_sobol = rnorm(mean = k, sd = k*sd, n = n_sample)
param_1 = cbind.data.frame(r = r_sobol, g = g_sobol, k = k_sobol)

## sample set 2
r_sobol = rnorm(mean = r, sd = r*sd, n = n_sample)
g_sobol = rnorm(mean = g, sd = g*sd, n = n_sample)
k_sobol = rnorm(mean = k, sd = k*sd, n = n_sample)
param_2 = cbind.data.frame(r = r_sobol, g = g_sobol, k = k_sobol)

# adjust the two parameter sets to remove negatives
param_1 = param_1 %>% map_df(pmax, 0)
param_2 = param_2 %>% map_df(pmax, 0)

# get sobol object from parameter sets
sens_c = sobolSalt(model = NULL, X1 = param_1, X2 = param_2, nboot = 300)

# rename columns
colnames(sens_c$X) = c("r", "g", "k")
```

```{r}
forest_growth_wrapper <- function(r, g, k, canopy_closure, C_initial, times, func) {
    # create a list of our parameters to work with the ODE solver
    params = list(r=r, g=g, k=k, canopy_closure=canopy_closure)
    
    # store the ODE results in a variable
    result = as.data.frame(ode(y=C_initial, times=times, func=func, parms=params))
    
    # rename columns and turn into df
    colnames(result) = c("time","C")
    
    # pull out the max forest size from our results df
    max_c = max(result$C)
    
    # return the max forest size as a list
  return(list(max_c = max_c))
}

# now use pmap to run our c_wrapper fxn over 
# a vector of parameter values from our sobol output
all_results = as.data.frame(sens_c$X) %>% 
  pmap(forest_growth_wrapper, canopy_closure = canopy_closure, C_initial = C_initial, 
       times = sim_length, func = forest_growth)

## extract values from bad list format
all_res = all_results %>% map_dfr(`[`,"max_c")
```


```{r}
# pivot table for plotting purposes
all_res_graph = all_res %>% 
  gather(key = "metric", value = "value")

# create the boxplot
ggplot(data = all_res_graph, aes(x = metric, y = value)) +
  theme_minimal() +
  geom_boxplot(fill = "#7E8F13", color = "#3E5409",
               alpha = 0.7, size = 0.7, width = 0.3) +
  labs(y = "Maximum Forest Size (kgC)\n",
       x = element_blank()) +
  theme(axis.title.y = element_text(size = 12, family = "Avenir", face = 'bold'),
        axis.text.x = element_blank()
  )
```

```{r, echo = FALSE}
# sobol indices
sens_forest = sensitivity::tell(sens_c, all_res$max_c)

# make a parameter column
parameter <- c("r","g","k")

# add the parameter column to make tables
sens_T <- cbind(parameter, as.data.frame(sens_forest$T))
rownames(sens_T) <- NULL

sens_S <- cbind(parameter, as.data.frame(sens_forest$S))
rownames(sens_S) <- NULL

# create tables for each index
sens_T %>% 
  kable(caption = "Table 1. First-order sobol indices") %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "striped",
                position = "left")

sens_S %>% 
  kable(caption = "Table 2. Total-effect indices") %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = "striped",
                position = "left")
```

Our first order sobol indices do not take covariance into account. Following this, we examined the total-effects indices, which DOES account for covariance. These results are helpful in explaining how our parameters interact and affect forest growth. In both cases, pre-canopy closure growth rate (r) and carrying capacity (K) have the biggest impact on forest growth, while post-canopy closure growth rate has a smaller impact on forest growth.

It's important to understand these interactions when managing a forest, especially in the face of climate change. 
